---
title: ' Crimes in Chicago - A Predictive Approach'
author: By Rushiraj Nenuji,Nischitha Krishna, Chandan Ramanna, Santhosh Kumar Ravindran
  and Yin Yang
date: "March 5, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
subtitle: Milestone 2 - Data Analysis I
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Regression models

## Loading data set

```{r}

crimedf <- read.csv(file = "Crimes_in_Chicago.csv", stringsAsFactors = FALSE)
```

## Cleansing data
```{r}

crimedf <- na.omit(crimedf)
```

## Convert Domestic into Dummy 1 for true and 0 for false


```{r}
crimedf$domestic_n <- ifelse (crimedf$Domestic == 'true',1,0)

```

# Regression models

## Convert Domestic into Dummy 1 for true and 0 for false


```{r}
crimedf$domestic_n <- ifelse (crimedf$Domestic == 'true',1,0)

```

## Filtering 1000 records for past2 years.
```{r}

crimedf1 <- subset(crimedf,Year == 2017)

library(magrittr)
library(dplyr)
Crimes_Type <- crimedf1 %>% group_by(Primary.Type) %>% dplyr:: summarise(arrest_count = n())  

Crimes_Type <- data.frame (Crimes_Type)
Crimes_Type<-Crimes_Type[order(-Crimes_Type$arrest_count),]
Crimes_Type <- head(Crimes_Type,5)
top10.crimes <- c(Crimes_Type$Primary.Type)

crimedf2 <- subset(crimedf1, Primary.Type %in% top10.crimes)

crimedf.new <- crimedf2[c(6:10,14,18)]
crimedf_filtered <- head(crimedf.new,1000)
```
## Logistic regression model

## Train a logistic regression model with 10-fold cross-validation
```{r}
library(caret)
fitControl <- trainControl(method = "cv",number = 15)
library(caret)
set.seed(123)
logit_fit <- train(Arrest ~ ., data = crimedf_filtered[1:7],
                   trControl = fitControl,
                   method="glm", family=binomial(link='logit'))

print(logit_fit)

confusionMatrix(logit_fit)
```
## Train Support Vector Machine (Radial Basis Function Kernel) with 2-fold Cross-Validation
```{r}
set.seed(123)
svmRadial_fit <- train(Arrest ~ ., data = crimedf_filtered[1:7],
                       trControl = fitControl, method = "svmRadial",
                       verbose=FALSE)

print(svmRadial_fit)

confusionMatrix(svmRadial_fit)
```

## Train Gradient Boosted Machine (GBM)
```{r}
library(plyr)
set.seed(123)
gbm_fit <-train(Arrest ~ ., data = crimedf_filtered[1:7],
                       trControl = fitControl, method = "gbm",
                       verbose=FALSE)
 
print(gbm_fit)

confusionMatrix(gbm_fit)
```
## Compare Different Predictive Models


## Data Partitioning

```{r}
crimedf_filtered$THEFT <- ifelse(crimedf_filtered$Primary.Type == 'THEFT', 1,0)
crimedf_filtered$BATTERY <- ifelse(crimedf_filtered$Primary.Type == 'BATTERY', 1,0)
crimedf_filtered$ASSAULT <- ifelse(crimedf_filtered$Primary.Type == 'ASSAULT', 1,0)
crimedf_filtered$CRIMINAL.DAMAGE <- ifelse(crimedf_filtered$Primary.Type == 'CRIMINAL DAMAGE', 1,0)
crimedf_filtered$OTHER.OFFENSE <- ifelse(crimedf_filtered$Primary.Type == 'OTHER OFFENSE', 1,0)
crimedf_filtered$Arrest1 <- ifelse(crimedf_filtered$Arrest == 'true', 1,0)
crimedf_filtered$Domestic1 <- ifelse(crimedf_filtered$Domestic == 'true', 1,0)
crimedf_filtered$Year <- as.numeric(crimedf_filtered$Year)
crimeneuraldf <- crimedf_filtered[c(7:14)]
crimeneuraldf$Year <- ifelse(crimedf_filtered$Year == 2017,1,0)
```

##Structure 
```{r}
str(crimeneuraldf)
```
## We use a single 80/20% split.

```{r}
set.seed(1234)
library(caret)
trainIndex <- createDataPartition(crimeneuraldf$Arrest1, p = .8, list = FALSE)
head(trainIndex)

train_data <- crimeneuraldf[ trainIndex,]
test_data  <- crimeneuraldf[-trainIndex,]
```

##Fit ANN Model on the Training Dataset

We use the neuralnet R package for neural network modeling. For a list of prediction and classification models, refer to http://topepo.github.io/caret/available-models.html

```{r}
# Load the neuralnet package
library('neuralnet')
```

After we load the neuralnet package, we can use the neuralnet() method to train the neural network model.

The 'hidden' is used to set a vector of integers specifying the number of hidden neurons (vertices) in each layer.

```{r}
f <- as.formula(Arrest1 ~ Year+THEFT+BATTERY+ASSAULT+CRIMINAL.DAMAGE+OTHER.OFFENSE+Domestic1)
```

# Fit a neural network model with 2 hidden layers
```{r}
nn_fit_2 <- neuralnet(f, data = train_data, hidden = c(5,3), linear.output=TRUE)
```

Now, let's show the structure of the neural network trained.

```{r}
plot(nn_fit_2,rep="best", cex=0.8)
```

# Show results

```{r}
summary(nn_fit_2)

```

# Evaluate Predictive Performance of the ANN Model 

## Computes the outputs of the ANN model on the test dataset

Notice that the prediction result of the neuralnet is the probability of belonging to a class rather than the class.

Hint:

- Use the compute() method to calculate prediction;
- In the compute() method, you need to NOT include the 'Personal_Loan' column, or else it will pop up an error of non-conformable arguments.  


```{r}
pred2_norm <- compute(nn_fit_2, test_data[-1])
pred2_norm <- pred2_norm$net.result

```


```{r}
test_data$pred <- ifelse(pred2_norm >= 0.2, 1, 0 )

```

## Create a confusion matrix to show the performance of the ANN model on the test set.


```{r}
confusionMatrix(test_data$Arrest1,test_data$pred)
```